Final Ratings for Genome Biology

### 2018-10-22 manuscript_figures/manuscript_ratings.Rmd
Add figures with tables for manuscript-ratings

```{r Figure2-load-data}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
##library(SummarizedBenchmark)
devtools::load_all("../../SummarizedBenchmark")

@@

# Also assumes that simulation results summary file is in the following location
# The `result-metrics.rds` was generated in the 
# `datasets/simulations/simulations-summary.Rmd` file.
resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

@@

```{r load-sim-metrics}
resmet <- readRDS(resmet_file)
```

@@

## read just real data samples
if (file.exists("datres.rds")) {
    datres <- readRDS("datres.rds")
} else {
    datres <- lapply(objects_dat, readRDS)
    saveRDS(datres, "datres.rds")
}

@@

## clean up names
datres <- dplyr::group_by(datres, casestudy, dataset, Method)
datres <- dplyr::summarize(datres, ntotal = n(), nNA = sum(is.na(adjp)),
                           nrej = sum(adjp < alpha.thresh, na.rm = TRUE))
datres <- dplyr::ungroup(datres)

@@

## add method rank in each category
datres <- dplyr::group_by(datres, casestudy, dataset)
datres <- dplyr::mutate(datres, rank = rank(-nrej)) 
datres <- dplyr::mutate(datres, rank = ifelse(is.na(nrej), NA, rank))
datres <- dplyr::ungroup(datres)

##saveRDS(datres, "datres-table.rds")
```

@@

We can use the results stored in `resmet` for simulation settings.
For easier plotting, we replace create a shorter distribution string column.

```{r}
resmet <- dplyr::mutate(resmet, sdist = factor(dist, sort(levels(dist)), c("chisq", "normal", "t11", "t5")))
```

We only look at results at the desired alpha cutoff, and resmet `fdrreg-e` from this comparison.

```{r}
resmet <- dplyr::filter(resmet, Method != "fdrreg-e", alpha == alpha.thresh)
```

@@

```{r}
resmet <- dplyr::group_by(resmet, setting, dist, inform, pi0, signal, ntests, sdist, key)
resmet <- dplyr::mutate(resmet, rank = rank(-mean.info))
resmet <- dplyr::ungroup(resmet)
```

@@

```{r}
resmet <- dplyr::mutate(resmet, setting = ifelse(setting == "varyingpi0-t", "varyingpi0", setting))
```

@@

```{r}
dplyr::filter(resmet, key == "FDR") %>%
    tidyr::unite(subsetting, sdist, inform, pi0, signal, ntests) %>%
    ggplot(aes(x = subsetting,
               y = mean.info, ymin = mean.info - se.info, ymax = mean.info + se.info, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting, space = "free_x", scales = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Simulations: FDR control across settings, alpha = ", alpha.thresh))

@@

```{r}
## FDR control across simulation settings
fdrcontrol_sims <- resmet %>%
    dplyr::filter(key == "FDR", !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean.info - se.info) > alpha.thresh),
                     pExceeds = round(nExceeds / n() * 100, 2)) %>%
    dplyr::ungroup()
dplyr::arrange(fdrcontrol_sims, pExceeds)

## FDR control across yeast settings
fdrcontrol_yeast <- yeast %>%
    dplyr::filter(key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean - se) > alpha.thresh),
                     pExceeds = round(nExceeds / n() * 100, 2)) %>%
    dplyr::ungroup()
dplyr::arrange(fdrcontrol_yeast, pExceeds)

@@

@@

```{r}
## power across simulation settings
power_sims <- resmet %>%
    dplyr::filter(key == "TPR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()
dplyr::arrange(power_sims, meanRank)

## power across yeast settings
power_yeast <- yeast %>%
    dplyr::filter(key == "TPR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()
dplyr::arrange(power_yeast, meanRank)

@@

@@

```{r}
## robustness as variance of FDR/TPR as % increase vs BH/qvalue
robustness_sims <- resmet %>%
    dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
                     sdGainBH = sd(mean.info / bh),
                     sdGainQ = sd(mean.info / qvalue),
                     sdGain = sd(mean.info / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()
dplyr::arrange(robustness_sims, key, -cvGain)

## robustness as variance of FDR/TPR as % increase vs BH/qvalue
robustness_yeast <- yeast %>%
    dplyr::filter(key == "TPR" | key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::select(Method, mean, key, setting_de, setting_cov, setting_mode, setting_pi0) %>%
    tidyr::spread(Method, mean) %>%
    tidyr::gather(Method, mean, -key, -setting_de, -setting_cov, -setting_pi0,
                  -setting_mode, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean / (bh + qvalue) * 2),
                     sdGainBH = sd(mean / bh),
                     sdGainQ = sd(mean / qvalue),
                     sdGain = sd(mean / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()
dplyr::arrange(robustness_yeast, key, -cvGain)

robustness_data <- datres %>%
    dplyr::mutate(prej = nrej / ntotal) %>%
    dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
    tidyr::spread(Method, prej) %>%
    tidyr::gather(Method, prej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     sdGainBH = sd(prej / bh, na.rm = TRUE),
                     sdGainQ = sd(prej / qvalue, na.rm = TRUE),
                     sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "PropRej")
dplyr::arrange(robustness_data, -cvGain)

@@

>>>>

```{r load-workspace, message = FALSE}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
##library(SummarizedBenchmark)
devtools::load_all("../../SummarizedBenchmark")

@@

# Also assumes that simulation results summary file is in the following location
# The `result-metrics.rds` was generated in the 
# `datasets/simulations/simulations-summary.Rmd` file.
simres_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

@@

```{r load-sim-metrics}
simres <- readRDS(simres_file)
```

@@

## read just real data samples
if (file.exists("datres.rds")) {
    datres <- readRDS("datres.rds")
} else {
    datres <- lapply(objects_dat, readRDS)

@@

    ## summarize results
    datres <- dplyr::group_by(datres, casestudy, dataset, Method)
    datres <- dplyr::summarize(datres, ntotal = n(), nNA = sum(is.na(adjp)),
                               nrej = sum(adjp < alpha.thresh, na.rm = TRUE))
    datres <- dplyr::ungroup(datres)

    saveRDS(datres, "datres.rds")
}

@@

## add method rank in each category
datres <- dplyr::group_by(datres, casestudy, dataset)
datres <- dplyr::mutate(datres, rank = rank(-nrej)) 
datres <- dplyr::mutate(datres, rank = ifelse(is.na(nrej), NA, rank))
datres <- dplyr::ungroup(datres)
```

@@

We can use the results stored in `simres` for simulation settings.
For easier plotting, we replace create a shorter distribution string column.

```{r}
simres <- dplyr::mutate(simres, sdist = factor(dist, sort(levels(dist)), c("chisq", "normal", "t11", "t5")))
```

We only look at results at the desired alpha cutoff, and remove `fdrreg-e` from this comparison.

```{r}
simres <- dplyr::filter(simres, Method != "fdrreg-e", Method != "bonf", alpha == alpha.thresh)
```

@@

```{r}
simres <- dplyr::group_by(simres, setting, dist, inform, pi0, signal, ntests, sdist, key)
simres <- dplyr::mutate(simres, rank = rank(-mean.info))
simres <- dplyr::ungroup(simres)
```

@@

```{r}
simres <- dplyr::mutate(simres, setting = ifelse(setting == "varyingpi0-t", "varyingpi0", setting))
```

@@

```{r}
dplyr::filter(simres, key == "FDR") %>%
    tidyr::unite(subsetting, sdist, inform, pi0, signal, ntests) %>%
    ggplot(aes(x = subsetting,
               y = mean.info, ymin = mean.info - se.info, ymax = mean.info + se.info, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting, space = "free_x", scales = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Simulations: FDR control across settings, alpha = ", alpha.thresh))

@@

```{r}
## FDR control across simulation settings
fdrcontrol_sims <- simres %>%
    dplyr::filter(key == "FDR", !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean.info - se.info) > alpha.thresh),
                     pExceeds = round(nExceeds / n(), 4)) %>%
    dplyr::ungroup()

## FDR control across yeast settings
fdrcontrol_yeast <- yeast %>%
    dplyr::filter(key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean - se) > alpha.thresh),
                     pExceeds = round(nExceeds / n(), 4)) %>%
    dplyr::ungroup()

@@

We plot the results as well.

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = fdrcontrol_yeast, sims = fdrcontrol_sims, .id = "analysis") %>%
    ggplot(aes(x = analysis, y = pExceeds, color = Method, group = Method)) +
    geom_line(aes(linetype = Method), alpha = 1/2, position = position_dodge(width = 1/4)) +
    geom_point(position = position_dodge(width = 1/4)) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Percent Settings Exceeding nominal FDR", labels = scales::percent) + 
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("FDR control across simulations and yeast experiments") 
```

@@

```{r}
## power across simulation settings
power_sims <- simres %>%
    dplyr::filter(key == "TPR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()

## power across yeast settings
power_yeast <- yeast %>%
    dplyr::filter(key == "TPR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()

@@

We plot the results as well.

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, meanRank)) %>%
    ggplot(aes(x = Method, y = meanRank, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("TPR rank across simulations and yeast experiments") 
```

@@

```{r}
## robustness across simulation settings
robustness_sims <- simres %>%
    dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
                     sdGainBH = sd(mean.info / bh),
                     sdGainQ = sd(mean.info / qvalue),
                     sdGain = sd(mean.info / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## robustness across yeast experiments
robustness_yeast <- yeast %>%
    dplyr::filter(key == "TPR" | key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::select(Method, mean, key, setting_de, setting_cov, setting_mode, setting_pi0) %>%
    tidyr::spread(Method, mean) %>%
    tidyr::gather(Method, mean, -key, -setting_de, -setting_cov, -setting_pi0,
                  -setting_mode, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean / (bh + qvalue) * 2),
                     sdGainBH = sd(mean / bh),
                     sdGainQ = sd(mean / qvalue),
                     sdGain = sd(mean / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## robustness across case studies
robustness_data <- datres %>%
    dplyr::mutate(prej = nrej / ntotal) %>%
    dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
    tidyr::spread(Method, prej) %>%
    tidyr::gather(Method, prej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     sdGainBH = sd(prej / bh, na.rm = TRUE),
                     sdGainQ = sd(prej / qvalue, na.rm = TRUE),
                     sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "PropRej")

@@

We plot the results as well.

```{r, fig.width = 6, fig.height = 8}
bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") + 
    expand_limits(y = 1) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of percent relative to BH/qvalue") + 
    expand_limits(y = 0) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of percent relative to BH/qvalue") +
    scale_linetype("Metric") + 
    expand_limits(y = 0) +
    facet_grid(analysis ~ ., scales = "free_y") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") +
    scale_linetype("Metric") + 
    expand_limits(y = 0) +
    facet_grid(analysis ~ ., scales = "free_y") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies") 
```
